Decade of Lost Precision:\\ A General Framework For Image Feature Matching Without Geometric Constraints

### 2014-04-16 text/journal/paper.tex
Modified paper to integrate stefan's changes

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Computer vision applications that involve the matching of local image 
features frequently use \emph{Ratio-Match} as introduced by Lowe and 
others, but is this really the optimal approach?  We formalize the 
theoretical foundation of \emph{Ratio-Match} and propose a general 
framework encompassing \emph{Ratio-Match} and five other matching 
methods. Using this framework, we establish a theoretical 
performance ranking in terms of precision and recall, 
proving that all five methods consistently outperform or equal \emph{Ratio-Match}.

@@

A plethora of solutions have combined ratio match with various geometric 
constraints to improve matching. These constraints are
based on assumptions regarding the transformation between the query and 
target images. At the stricter end we have epipolar constraints, assuming
that the images can be tied by a homography \cite{torr2000mlesac,chum2005matching}, and angular constraints, assuming the 
correspondences are angled similarly \cite{kim2008efficient,schmid1997local,leordeanu2005spectral}. 
These constraints can be relaxed using more sophisticated models.  Torresani et al.\ \cite{torresani2008feature} 
attempted to model transformations using graph matching similar to 
\cite{deriche1994robust}. Others define image regions and reject or 
accept correspondences based on the regions they connect 
\cite{cho2009feature,wu2011robust}.

@@

The paper is organized as follows. In Section~\ref{S:MatchingMethods} we 
introduce the original \emph{Ratio-Match}, generalize this method to 
introduce the proposed framework, and compare the methods of the 
framework theoretically.  In Section~\ref{S:Experiments} we present 
our experimental validation and discuss the results obtained.  Section~\ref{S:Summary} 
concludes the paper.

\section{Matching Framework}
\label{S:MatchingMethods}
%
\subsection{Ratio-Match}
%
\emph{Ratio-Match} as introduced by Deriche et al.\ 
\cite{deriche1994robust} and later used by 
Baumberg~\cite{baumberg2000reliable} and Lowe~\cite{lowe2004sift} makes 
use of a \emph{uniqueness ratio}. For each feature in the \emph{query 
image}, it finds a possible match by identifying the nearest neighbor and 
returning this match only if the \emph{uniqueness ratio} is above a 
given threshold $\tau$.

Given $f_q$, a feature from the \emph{query 
image} and $F_{target}$, a set of features from the \emph{target image}, 
the nearest neighbor of $f_q$ in $F_{target}$ is:
\begin{equation*}
    f_{nn} = \argmin_{f_t \in F_{target}} d(f_q, f_t).
\end{equation*}
Here, $d(f_i, f_j)$ is the distance between feature descriptors.  
With SIFT and SURF this is the Euclidean distance of the feature vectors
\cite{lowe2004sift,bay2006surf}, whereas BRIEF, BRISK, and FREAK 
use the Hamming distance \cite{leutenegger2011brisk,calonder2010brief,alahi2012freak}.

If we let $f_p$ and $f_b$ be the nearest and second nearest neighbor of a feature $f_q$, 
the \emph{uniqueness ratio} is defined as follows:
\begin{equation*}
    r = d(f_q, f_p) / d(f_q, f_b).
\end{equation*}
Because $f_p$ is the nearest neighbor and $f_b$ the second nearest 
neighbor, we know that $d(f_q, f_p) \leq d(f_q, f_b)$, and since $d(f_i, 
f_j) \geq 0, \forall f_i,f_j$, the \emph{uniqueness ratio} is bounded by $0$ 
and $1$, both inclusive. For the degenerate case where $d(f_q, f_b) = 0$
we let the \emph{uniqueness ratio} be $0$.

The underlying assumption in ratio match is that the point we seek to 
match in the \emph{query image} has only one match in a given 
\emph{target image} or no matches at all. In both cases we should expect 
the \emph{second} nearest neighbor in the target image to be an 
untrue correspondence.  This means we can use the distance between the 
feature point and the second closest descriptor as a \emph{baseline} for 
false correspondences. By dividing the distance to the nearest neighbor 
with this baseline, we get an estimated measure for how distinct the 
correspondence is from a false match.

To make further discussion easier we introduce the following 
nomenclature:
\begin{itemize}
\item{Let $f_q$ be any feature point in the \emph{query image} that we 
    are currently trying to match}.
\item{Let $F$ be a set of features. $F_{target}$ for example is the set 
    of features from the \emph{target image}}.
\item{Let $\tau \in [0 \ldots 1]$ be a threshold used to decide whether 
	to keep a match or discard it. We keep a match only if it has a 
uniqueness ratio lower than $\tau$}.
\item{Let the \emph{proposed match} be the nearest neighbor of $f_q$
    picked from a set of feature points that we will call the 
\emph{proposal set} which does not include $f_q$}.
\item{Let the \emph{baseline match} be the nearest neighbor of $f_q$ 
    picked from a set of feature points that we will call the 
\emph{baseline set}, which neither contains the \emph{proposed match} nor 
$f_q$}.
\end{itemize}

In the case of \emph{Ratio-Match} \cite{lowe2004sift}, the \emph{proposal 
set} consists of all the feature points in the target image.  Similarly 
the \emph{baseline set} consists of all the feature points in the target 
image except the \emph{proposed} match. For \emph{Ratio-Match} this 
practically means that given a feature $f_q$, the \emph{proposed match} 
and the \emph{baseline match} are the two nearest neighbors in the set 
of feature points in the \emph{target image}.

Brown et al.\ \cite{brown2005multi} extend \emph{Ratio-Match} for the case of $n$ target 
images by letting the baseline be the average of the distance to the 
second nearest neighbors for all $n$ target images.  They write, ``\emph{[W]e can improve outlier rejection by using information from all of the images (rather 
than just the two being matched). This improves the matching precision}''.  The idea of improving matching precision by expanding the 
knowledge used to reject outliers is also the central tenet of the framework presented in this paper.

%
\subsection{Framework of Matching Methods}
%

\emph{Ratio-Match} is only one instance of a framework of six algorithms we can 
construct by different combinations of \emph{proposed} and 
\emph{baseline sets}. As shown in Table~\ref{table:distinctness} for the 
case of two images, the \emph{proposal set} can be constructed by either 
looking only at the nearest neighbors in the target image, or 
alternatively by looking at the nearest neighbors in both the query and 
target image. The \emph{baseline set} can either be the query image 
itself, the target image, or the combination of the two.  Seen in this 
light, the matching algorithm \emph{Ratio-Match} draws both its 
\emph{proposal set} and \emph{baseline set} from the target image. In 
turn \emph{Mirror-Match-Ext} as introduced by the authors in 
\cite{arnfred2013mirror}, works by constructing the \emph{proposal set} 
and \emph{baseline set} by combining features from both the \emph{query} 
and \emph{target image}. Alternatively \emph{Self-Match} looks only to 
the query image for the \emph{baseline set} but constructs the 
\emph{proposal set} from the target image like \emph{Ratio-Match}. In 
Figure~\ref{fig:overview} we illustrate how each method is derived.

\begin{table}[htb]
\caption{Classification of the methods of the proposed framework based on the feature sources for 
    \emph{proposal set} and \emph{baseline set} for calculating the \emph{uniqueness ratio}.}
\label{table:distinctness}
	\centering
\bgroup
\def\arraystretch{1.5}
    \begin{tabular}{l l | l l}
        %
        %&  \multicolumn{1}{l}{Target}
        %& \multicolumn{1}{l}{Target and Query} \\
\multicolumn{2}{r}{} & \multicolumn{2}{c}{\emph{Proposal Set}} \\
                     & & Target & Target and Query \\
        \cline{2-4}
        \multirow{3}{*}{\rotatebox[origin=c]{90}{\emph{Baseline Set}}} &
  Query            & \emph{Self-Match}   & \emph{Self-Match-Ext} \\
& Target           & \emph{Ratio-Match}   & \emph{Ratio-Match-Ext} \\
& Target and Query & \emph{Mirror-Match} & \emph{Mirror-Match-Ext} \\
        \cline{2-4}
\end{tabular}
\egroup
\end{table}

\begin{figure}[htb]
\centering
\includegraphics[width=0.85\columnwidth]{images/overview}
\caption{Graphical representation of the \emph{baseline} and 
\emph{proposal sets} for the methods in the proposed framework.}
\label{fig:overview}
\end{figure}

Algorithm~\ref{alg-gen} shows the generalized matching framework, 
where the \emph{proposal set} and \emph{baseline set} are defined 
according to Table~\ref{table:distinctness}.  In practice for e.g.\ 
\emph{Ratio-Match} we can obtain $f_p$ and $f_b$ by finding the two 
nearest neighbors of $F_{target}$. The other algorithms can be optimized 
similarly. Figure~\ref{fig:matching} illustrates the structure of 
the generalized algorithm.

\begin{algorithm}[htb]
\caption{Generalized matching algorithm for two images}
\label{alg-gen}
\begin{algorithmic}
    \Require $I_{query}, I_{target}$ : images, $\tau \in [0,1]$
    \State $F_{query} = get\_features(I_{query})$
    \State $F_{target} = get\_features(I_{target})$
\State $F_{proposal} = get\_proposal\_set(F_{query}, F_{target})$
\State $F_{baseline} = get\_baseline\_set(F_{query}, F_{target})$
\State $M = \varnothing$
\ForAll{$f_q \in F_{query}$}
    \State $f_p \gets getNearestNeighbo\text{r}(f_q, F_{proposal} \setminus 
    \left\{f_q\right\})$
    \State $f_b \gets getNearestNeighbo\text{r}(f_q, F_{baseline} \setminus 
    \left\{f_q, f_p\right\})$
    \State $ratio \gets distance(f_q, f_p) / distance(f_q, f_b)$
    \If{$(ratio < \tau) \wedge (f_p \in F_{target})$}
        \State $matches \gets matches \cup \left(f_q, f_p\right)$
	\EndIf
\EndFor
\Return $M$
\end{algorithmic}
\end{algorithm}

@@

For some methods it is possible that the \emph{proposal match} is a 
feature from the \emph{query image}, in which case we discard the match 
as an untrue correspondence.  It also happens that we encounter 
correspondences with a \emph{uniqueness ratio} above $1$, in which case we set 
it to $1$ to maintain a bounded \emph{uniqueness ratio}.  Take for example the 
case of \emph{Self-Match}. Using this method we might find that the 
nearest neighbor of a feature in the target image is further from the 
query feature than the nearest neighbor in the query image. In this case 
the match is discarded.

\subsection{Generalized Uniqueness Ratio}

Form $F_{query} = \left\{f_{q_1} \ldots f_{q_n}\right\}$ and $F_{target} 
= \left\{f_{t_1} \ldots f_{t_m}\right\}$.  We can obtain $F_{proposed}$ 
and $F_{baseline}$ from these sets depending on the algorithm used (see 
Table~\ref{table:distinctness}). $r = \text{r}(f_{q}, F_{proposed}, 
F_{baseline})$ is the ratio of query feature $f_{q}$ given 
$F_{proposed}$ and $F_{baseline}$.  If $f_{p}$ is the nearest neighbor 
of $f_{q}$ in $F_{proposed}$ and $f_{b}$ is the nearest neighbor of 
$f_{q}$ in $F_{baseline} \setminus \left\{f_{p}\right\}$, then the 
generalized \emph{uniqueness} ratio is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \frac{d(f_{q}, f_{p})}{d(f_{q}, f_{b})}.
\end{align*}

We let $K$ be the number of possible true correspondences between \emph{query} 
and \emph{target image}. \emph{Precision} and \emph{recall} are 
defined as:
\begin{align*}
    \textrm{Recall} &= \frac{\#\textrm{Correct Matches}}{K} \\
    \textrm{Precision} &= \frac{\#\textrm{Correct 
    Matches}}{\#\textrm{Correct Matches} + \#\textrm{Incorrect Matches}}
\end{align*}
Given $f_{nn} \in F_{target}$ as the nearest neighbor of any 
$f_q \in F_{query}$, we define the set of features $F_{true}$ as 
$\left\{ f_{q} \in F_{query} \mid f_{nn} \text{ is a true correspondence 
of } f_{q} \right\}$ and $F_{false}$ as $F_{query} \setminus F_{true}$.  
We can then define the number of correct and incorrect matches as 
follows:
\begin{align*}
    \textrm{keep\_match}(f_{q}) &= \twopartdef{ 1 }{\text{r}(f_{q}, 
    F_{proposed}, F_{baseline}) <
    \tau}{0}{\textrm{otherwise}} \\
    \#\textrm{Correct Matches} &= \sum_{f_{q} \in F_{true}} 
    \textrm{keep\_match}(f_{q})\\
    %\mathbbm{1}_{\left\{ \right\}} \\
    \#\textrm{Incorrect Matches} &= \sum_{f_{q} \in F_{false}}
    \textrm{keep\_match}(f_{q})
    %\mathbbm{1}_{\left\{ r < \tau\right\}}
\end{align*}

From the above we can conclude that the performance in terms of 
precision and recall of any method from the proposed framework is 
uniquely identified by the \emph{uniqueness ratio} $r$.  If for every query 
feature $f_{q}$, $r$ is identical across two matching methods, then they 
return identical results.

@@

Under the following three assumptions, which largely reflect the performance of 
matching feature points in practice, we can theoretically compare the 
performance of the different algorithms shown in 
Table~\ref{table:distinctness}:
\begin{enumerate}
    \item{For a query and target image we assume that for any point in 
        the query image there is at most one real correspondence in the 
    target image and no real correspondences in the query image itself 
as assumed by \emph{Ratio-Match}.}
    \item{We assume that distances of feature descriptors are well 
            behaved.  More precisely, given $f_q$, a feature from the 
            \emph{query image}, and $f_{match}$, the real correspondence 
            from the set of features in the \emph{target image} 
            $F_{target}$, we have $\forall f_i \in F_{proposal}, 
            f_i \neq f_{match}: d(f_q,f_{match}) < d(f_q, f_i)$.}
    %\item{We assume that features within one image resemble each other 
            %more than features in different images. That is, given 
            %$f_a,
    %        f_i \in F_1$ and $f_b, f_j \in F_2$ where $(f_i, f_j)$ and 
    %    $(f_a, f_b)$ are true correspondences, and $F_1$ and $F_2$ are 
    %features collected from different images then we expect that 
            %$d(f_a, f_i)% < d(f_a, f_j)$ and $d(f_b, f_j) < d(f_b, 
            %f_i)$}
     \item{For a set of query features $F_{query}$ with a true 
             correspondence in the \emph{target image}, we assume that 
             the distribution of \emph{uniqueness ratios} using $F_{target}$ as 
             the baseline set is similar to using $F_{query}$, since the 
         two images in the case of true correspondences are guaranteed 
     to share part of the same scene.}
    \end{enumerate}

Based on these assumptions we first prove that the results of 
\emph{Self-Match} are identical to \emph{Self-Match-Ext}. To do this, we 
look at the nearest neighbor $f_{nn}$ of every query feature $f_{q}$ and 
consider the two cases of $f_{nn} \in F_{query}$ and $f_{nn} \in
F_{target}$. For $f_{nn} \in F_{target}$ the \emph{uniqueness ratio} of 
\emph{Self-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline}).
\end{align*}
Since $F_{proposed} = F_{target}$ for \emph{Self-Match}, this shows that
when the nearest neighbor is found in the target image, the two 
algorithms behave identically. For $f_{nn} \in F_{query}$ 
\emph{Self-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{target}, F_{query}) \\
        &\geq 1.
\end{align*}
The closest neighbor is found in the query image, so $d(f_{q}, f_{nn}) < 
d(f_{q},f_{p})$, where $f_{p}$ is the nearest neighbor from 
$F_{proposal}$. \emph{Self-Match-Ext} does not return a correspondence 
found in the \emph{query image}, and since $\tau \in [0,1]$, neither does 
\emph{Self-Match}. This proves that 
\emph{Self-Match} returns the same matches as \emph{Self-Match-Ext} for any set of query features. For 
this reason both algorithms will be referred to as just 
\emph{Self-Match}.

The above proof can easily be extended to show that \emph{Mirror-Match} and \emph{Mirror-Match-Ext} also return the same correspondences, and 
both will be referred to as \emph{Mirror Match} from here on.

For the case of \emph{Ratio-Match} and \emph{Ratio-Match-Ext}, we 
can show that \emph{Ratio-Match-Ext} is equal or better than 
\emph{Ratio-Match}.  Consider again the nearest neighbor $f_{nn}$ of a 
query feature $f_{q}$, and the two cases where we either have $f_{nn} 
\in F_{query}$ or $f_{nn} \in F_{target}$. For $f_{nn} \in F_{target}$ 
the \emph{uniqueness ratio} of \emph{Ratio-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline}).
\end{align*}
Since $F_{proposed} = F_{target}$ for \emph{Ratio-Match} this shows that
when the nearest neighbor is found in the target image, the two 
algorithms behave identically. For $f_{nn} \in F_{query}$ 
\emph{Ratio-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{target}, F_{target}).
\end{align*}
While \emph{Ratio-Match-Ext} does not return the correspondence if 
the nearest neighbor of a query feature is in the \emph{query image}, 
\emph{Ratio-Match} is likely to do so, since the \emph{uniqueness ratio} is 
found by dividing with the second nearest neighbor in the \emph{target 
image}. This shows that for cases where a query feature has a nearest 
neighbor in the \emph{query image}, \emph{Ratio-Match-Ext} is superior 
to \emph{Ratio-Match}. Since query features with a nearest neighbor in 
the \emph{query image} are false correspondences per assumptions \#1 and 
\#2, this proves that \emph{Ratio-Match-Ext} has superior 
\emph{Precision} to \emph{Ratio-Match} while maintaining equal 
\emph{Recall}.

Finally we show that the accuracy of \emph{Mirror-Match} is equal to or better than 
\emph{Ratio-Match-Ext}. Consider as before the nearest neighbor $f_{nn}$ 
of a query feature $f_{q}$. When $f_{nn}$ resides in the \emph{query 
image}, both algorithms behave alike and discard the match.  However, 
consider the \emph{uniqueness ratio} of \emph{Mirror Match} for the case where 
$f_{nn}$ resides in the \emph{target image}:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{proposed}, F_{query} \cup F_{target})\\
        &= max( \text{r}(f_{q}, F_{proposed}, F_{target}), 
    \text{r}(f_{q}, F_{proposed}, F_{query}) ).
\end{align*}
For a true correspondence the \emph{uniqueness ratio} using 
$F_{query}$ as a baseline is distributed similarly to the ratio when using 
$F_{target}$ according to assumption \#3. In this case the algorithm performs like 
\emph{Ratio-Match-Ext}.  However, for an untrue correspondence with a 
baseline match in the \emph{query image} which is closer than the
baseline match in the \emph{target image}, \emph{Mirror Match} will 
return a worse \emph{uniqueness ratio}. This in turns means that \emph{Mirror 
Match} is equal or better than \emph{Ratio-Match-Ext} in terms of  
\emph{Precision} while maintaining equal \emph{Recall}.

We have thus proven that -- based on assumptions \#1-3 -- for any given recall rate, the 
precision of the algorithms in the framework compares as follows:
\begin{align*}
    \textit{Self-Match} &= \textit{Self-Match-Ext} \\
    \textit{Mirror-Match} &= \textit{Mirror-Match-Ext} \\
    \textit{Ratio-Match} &\leq \textit{Ratio-Match-Ext} \leq 
    \textit{Mirror Match}
\end{align*}

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Computer vision applications that involve the matching of local image features frequently use \emph{Ratio-Match} as introduced by Lowe and others, but is this really the optimal approach?  We formalize the theoretical foundation of \emph{Ratio-Match} and propose a general framework encompassing \emph{Ratio-Match} and three other matching methods. Using this framework, we establish a theoretical performance ranking in terms of precision and recall, proving that all five methods consistently outperform or equal \emph{Ratio-Match}.

@@

A plethora of solutions have combined ratio match with various geometric constraints to improve matching. These constraints are based on assumptions regarding the transformation between the query and target images. At the stricter end we have epipolar constraints assuming that the images can be tied by a homography \cite{torr2000mlesac, chum2005matching} and angular constraints assuming the correspondences are angled similarly \cite{kim2008efficient, schmid1997local}. Often these approaches are made computationally feasible by modelling the case of feature correspondences as an instance of graph matching where each feature is a vertex and edge values correspond to a geometric relation between two features. In this case approximate graph matching algorithms can be used to efficiently establish an isomorphism between the graph of features in two images \cite{leordeanu2005spectral, torresani2008feature, yarkony2010covering}. Finally others define image regions and reject or accept correspondences based on the regions they connect \cite{cho2009feature, wu2011robust}.

@@

In the author's last paper we introduced \emph{Mirror-Match} and \emph{Mirror-Match with Clustering}, two algorithms that we showed to perform better than the state of the art. The novel contributions of this paper consists of two new algorithms completing a framework of four related matching algorithms as well as the theoretical foundation which enables us to understand why \emph{Mirror-Match} performs better than \emph{Ratio Match} in the first place.

The paper is organized as follows. I section~\ref{S:MatchingMethods} we introduce the original \emph{Ratio-Match}, extend on this method to introduce the proposed framework and go on to compare the members of the framework theoretically.  In section~\ref{S:Experiments} we present evaluations and discuss the results obtained.  Section~\ref{S:Summary} concludes the paper.

\section{Matching Framework}
\label{S:MatchingMethods}
%
\subsection{Definitions}
%
The proposed framework is inspired by \emph{Ratio-Match} as introduced by Deriche et al.\ \cite{deriche1994robust} and later used by Baumberg~\cite{baumberg2000reliable} and Lowe~\cite{lowe2004sift}. \emph{Ratio-Match} makes use of a \emph{uniqueness ratio} as a heuristic for finding correspondences between feature points. For each feature in the \emph{query image}, it finds a possible match by identifying the nearest and second nearest neighbor and returns this match only if the ratio between the distance of the nearest and second nearest neighbor is above a given threshold $\tau$.

The underlying assumption in ratio match is that the point we seek to match in the \emph{query image} has only one match in a given \emph{target image} or no matches at all. In both cases we should expect the \emph{second} nearest neighbor in the target image to be an untrue correspondence.  This means that we can use the distance between the feature point and the second closest descriptor as a \emph{baseline} for false correspondences. By dividing the distance to the nearest neighbor with this baseline, we get an estimated measure for how distinct the correspondence is from a false match.

To make further discussion easier we introduce the following nomenclature:
\begin{itemize}
\item{Let $f_q$ be any feature point in the \emph{query image} that we 
    are currently trying to match}.
\item{Let $F$ be a set of features. $F_{target}$ for example is the set 
    of features from the \emph{target image}}.
\item{Let $\tau \in [0 \ldots 1]$ be a threshold used to decide whether 
    to keep a match or discard it. We keep a match only if is has a 
\emph{uniqueness ratio} lower than $\tau$}
\item{Let the \emph{proposed match} be the nearest neighbor of the query 
    feature $f_q$ picked from a set of feature points that we will call 
the \emph{proposal set} which does not contain the query feature itself}
\item{Let the \emph{baseline match} be the nearest neighbor of the query 
    feature $f_q$ picked from a set of feature points that we will call 
the \emph{baseline set} which neither contains the \emph{proposed match} 
nor the query feature}
\end{itemize}

%
\subsection{A framework of matching methods}
%

\begin{figure}[t]
\centering
\includegraphics[width=0.85\columnwidth]{images/overview}
\caption{Graphical representation of the \emph{Baseline} and 
\emph{Proposal Set} for different methods in the proposed framework}
\label{fig:overview}
\end{figure}

Given two images we can construct the proposal and baseline sets in six different ways as illustrated in Figure~\ref{fig:overview}. In the case of \emph{Ratio-Match} \cite{lowe2004sift} the \emph{proposal set} consists of all the feature points in the target image.  Similarly the \emph{baseline set} consists of all the feature points in the target image except the \emph{proposed} match. For \emph{Ratio-Match} this practically means that given a query feature, $f_q$, the \emph{proposed match} and the \emph{baseline match} are the two nearest neighbors in the set of feature points in the \emph{target image} and the ratio is decided uniquely based on the \emph{target image}. In turn \emph{Mirror-Match} as introduced by the authors in \cite{arnfred2013mirror}, works by constructing the \emph{proposal set} and \emph{baseline set} by combining features from both the \emph{query} and \emph{target image}. By similar variations of the \emph{proposal} and \emph{baseline set} we derive four more algorithms as illustrated in Figure~\ref{fig:overview}, however we will not discuss \emph{Self-Match-Ext} and \emph{Both-Match} further in this paper since we show them to be equivalent with \emph{Self-Match} and \emph{Mirror-Match} respectively in Appendix~\ref{A:self} and \ref{A:mirror}.

\begin{algorithm}[htb]
\caption{Generalized matching algorithm for two images}
\label{alg-gen}
\begin{algorithmic}
    \Require $I_{query}, I_{target}$ : images, $\tau \in [0,1]$
    \State $F_{query} = get\_features(I_{query})$
    \State $F_{target} = get\_features(I_{target})$
\State $F_{proposal-all} = get\_proposal\_features(F_{query}, 
F_{target})$
\State $F_{baseline-all} = get\_baseline\_features(F_{query}, 
F_{target})$
\State $M = \varnothing$
\ForAll{$f_q \in F_{query}$}
    \State $F_{proposal} = F_{proposal-all} \setminus 
    \left\{f_q\right\}$
    \State $f_p \gets getNearestNeighbo\text{r}(f_q, F_{proposal})$
    \State $F_{baseline} = F_{baseline-all} \setminus \left\{f_q, 
    f_p\right\}$
    \State $f_b \gets getNearestNeighbo\text{r}(f_q, F_{baseline})$
    \State $ratio \gets distance(f_q, f_p) / distance(f_q, f_b)$
    \If{$(ratio < \tau) \wedge (f_p \in F_{target})$}
        \State $matches \gets matches \cup \left(f_q, f_p\right)$
	\EndIf
\EndFor \\
\Return $M$
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg-gen} shows the generalized matching framework where the \emph{proposal set} and \emph{baseline set} are defined according to Figure~\ref{fig:overview}.  In practice for e.g.\ \emph{Ratio-Match} we can obtain $f_p$ and $f_b$ by finding the two nearest neighbors of $F_{target}$. The other algorithms can be optimized similarly. Figure~\ref{fig:matching} illustrates the structure of this algorithm.

@@

For some methods it is possible that the \emph{proposal match} is a feature from the \emph{query image}, in which case we discard the match as an untrue correspondence.  It also happens that we encounter correspondences with a \emph{uniqueness ratio} above $1$, in which case the match is also discarded.  Take for example the case of \emph{Self-Match} where we might find that the nearest neighbor of a feature in the target image is further from the query feature than the nearest neighbor in the query image. In this case the match is discarded.

%
\subsection{Uniqueness Ratio}
%

The main distinguishing factor in between the algorithms in the framework is the final ratio between the distance of the two nearest neighbors of a query feature which decides if we keep or discard a match. We illustrate this process in Figure~\ref{fig:matching}.

We can define a calculation of this ratio common to all algorithms in the framework which allows us to compare the algorithms theoretically. To do so we first define the concept of a nearest neighbor and then based on that the \emph{uniqueness ratio}, $r$:

Given a feature, $f_i$ and a set of features, $F$ the nearest neighbor of $f_i$ in $F$ is: 
\begin{equation*}
\argmin_{f_j \in F} d(f, f_j)
\end{equation*}
Here, $d(f_i, f_j)$ is the distance between feature descriptors.With SIFT and SURF this is the Euclidean distance of the feature vectors \cite{lowe2004sift,bay2006surf}, whereas BRIEF, BRISK, and FREAK use the Hamming distance \cite{leutenegger2011brisk,calonder2010brief,alahi2012freak}.

If we let $f_p \in F_{proposed}$ and $f_b \in F_{baseline}$ be the nearest neighbors of a feature $f_q$ in $F_{proposed}$ and $F_{baseline}$, the \emph{uniqueness ratio} is defined as follows:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \frac{d(f_{q}, f_{p})}{d(f_{q}, f_{b})}.
\end{align*}

The performance in terms of precision and recall of any algorithm from the proposed framework is uniquely identified by the \emph{uniqueness ratio} $r$. To show this we let $K$ be the number of possible true correspondences between \emph{query} and \emph{target image}. \emph{Precision} and \emph{recall} are defined as:
\begin{align*}
    \textrm{Recall} &= \frac{\#\textrm{Correct}}{K} \\
    \textrm{Precision} &= \frac{\#\textrm{Correct}}{\#\textrm{Correct} + \#\textrm{Incorrect}}
\end{align*}
Given $f_{p} \in F_{proposed}$ as the nearest neighbor of some $f_q \in F_{query}$, we define the set of features $F_{true}$ as $\left\{ f_{q} \in F_{query} \mid f_{p} \text{ is a true correspondence of } f_{q} \right\}$ and $F_{false}$ as $F_{query} \setminus F_{true}$.  We can then define the number of correct and incorrect matches as follows:
\begin{align*}
    \textrm{keep\_match}(f_{q}) &= \twopartdef{ 1 }{\text{r}(f_{q}, 
    F_{proposed}, F_{baseline}) <
    \tau}{0}{\textrm{otherwise}} \\
    \#\textrm{Correct} &= \sum_{f_{q} \in F_{true}} 
    \textrm{keep\_match}(f_{q})\\
    %\mathbbm{1}_{\left\{ \right\}} \\
    \#\textrm{Incorrect} &= \sum_{f_{q} \in F_{false}}
    \textrm{keep\_match}(f_{q})
    %\mathbbm{1}_{\left\{ r < \tau\right\}}
\end{align*}
If for every query feature $f_{q}$, $r$ is identical across two matching methods, then they return identical results.

@@

Under the following three assumptions, which largely reflect the performance of matching feature points in practice, we can theoretically compare the performance of the different algorithms shown in Figure~\ref{fig:overview}:

\begin{enumerate}
    \item{For a query and target image we assume that for any point in 
        the query image there is at most one real correspondence in the 
    target image and no real correspondences in the query image itself 
as assumed by \emph{Ratio-Match}.}
    \item{We assume that distances of feature descriptors are well 
            behaved.  More precisely, given $f_q$, a feature from the 
            \emph{query image}, and $f_{match}$, the real correspondence 
            from the set of features in the \emph{target image} 
            $F_{target}$, we have $\forall f_i \in F_{proposal}, 
            f_i \neq f_{match}: d(f_q,f_{match}) < d(f_q, f_i)$.}
    %\item{We assume that features within one image resemble each other 
            %more than features in different images. That is, given 
            %$f_a,
    %        f_i \in F_1$ and $f_b, f_j \in F_2$ where $(f_i, f_j)$ and 
    %    $(f_a, f_b)$ are true correspondences, and $F_1$ and $F_2$ are 
    %features collected from different images then we expect that 
            %$d(f_a, f_i)% < d(f_a, f_j)$ and $d(f_b, f_j) < d(f_b, 
            %f_i)$}
     \item{For a set of query features $F_{query}$ with a true 
             correspondence in the \emph{target image}, we assume that 
             the distribution of \emph{uniqueness ratios} using $F_{target}$ as 
             the baseline set is similar to using $F_{query}$, since the 
         two images in the case of true correspondences are guaranteed 
     to share part of the same scene.}
    \end{enumerate}

Based on these assumptions we prove that \emph{Ratio-Match-Ext} is equal or better than \emph{Ratio-Match} in terms both precision and recall.  Consider the nearest neighbor $f_{p}$ of a query feature $f_{q}$ and the two cases where we either have the $f_{p} \in F_{query}$ or $f_{p} \in F_{target}$. For $f_{p} \in F_{target}$ the uniqueness ratio of \emph{Ratio-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline})
\end{align*}
Since $F_{proposed} = F_{target}$ for \emph{Ratio-Match} this shows that when the nearest neighbor is found in the target image, the two algorithms behave identically. For $f_{p} \in F_{query}$ \emph{Ratio-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{target}, F_{target}).
\end{align*}

That is, \emph{Ratio-Match} calculates the ratio based on the two nearest correspondences in the \emph{target image}. \emph{Ratio-Match-Ext} on the other hand does not return any correspondence because the nearest neighbor is in the \emph{query image}. Since query features with a nearest neighbor in the \emph{query image} are false correspondences per assumption \#1 and \#2, this proves that \emph{Ratio-Match-Ext} has superior \emph{Precision} to \emph{Ratio-Match} while maintaining equal \emph{Recall}.

Next we show that \emph{Mirror-Match} is equal or better than \emph{Ratio-Match-Ext} in terms of precision and recall. Consider as before the nearest neighbor $f_{p}$ of a query feature $f_{q}$. When $f_{p}$ resides in the \emph{query image} both algorithms behave alike and discard the match.  However, consider the uniqueness ratio of \emph{Mirror-Match} for the case where $f_{p}$ resides in the \emph{target image}:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{proposed}, F_{query} \cup F_{target})\\
        &= max( \text{r}(f_{q}, F_{proposed}, F_{target}), 
    \text{r}(f_{q}, F_{proposed}, F_{query}) ).
\end{align*}
For a true correspondence the \emph{uniqueness ratio} using $F_{query}$ as a baseline is distributed similarly to the ratio when using $F_{target}$ according to assumption \#3. In this case the algorithm performs like \emph{Ratio-Match-Ext}.  However, for an untrue correspondence with a baseline match in the \emph{query image} which is closer than the baseline match in the \emph{target image}, \emph{Mirror Match} will return a worse \emph{uniqueness ratio}. This in turns means that \emph{Mirror Match} is equal or better than \emph{Ratio-Match-Ext} in terms of  \emph{Precision} while maintaining equal \emph{Recall}.

Using a similar procedure we prove that \emph{Self-Match} is equal to \emph{Self-Match-Ext} and \emph{Both-Match} is equal to \emph{Mirror Match} in Appendix~\ref{A:self} and \ref{A:mirror} respectively.

We have thus proven that -- based on assumptions \#1-3 -- for any given recall rate, the precision of the algorithms in the framework compares as follows: 
\begin{align*}
    \textit{Ratio-Match} &\leq \textit{Ratio-Match-Ext} \leq 
    \textit{Mirror-Match}
\end{align*}
